---
title: "Lab 6 Comparing two means"
author: ""
output: github_document
---

### Run package and load data

```{r}
library(tidyverse)
fish <- read_csv("chap12q19ElectricFish.csv")
crabs <- read_csv("chap15q27FiddlerCrabFans.csv")
```

Researchers studying the number of electric fish species living in various parts of the Amazon basin were interested in whether the presence of tributaries affected the local number of electric fish species in the main rivers (Fernandes et al. 2004).

They counted the number of electric fish species above and below the entrance point of a major tributary at 12 different river locations.

The data is provided in your GitHub repository.

For each question below, write a sentence answering the question and show the code you used to come up with the answer, if applicable.

### Question A

> What is the mean different in the number of species between areas upstream and downstream of a tributary? What is the 95% confidence interval of this mean difference. Show your code and write a sentence giving your answer

ANSWER

```{r}
fish_long <- 
  pivot_longer(fish, speciesUpstream:speciesDownstream,
               names_to = "location",
               values_to = "species") %>% 
  mutate(location = str_remove(location, c("species"))) %>% 
  print()
t.test(formula = species ~ location, data = fish_long)
```

The mean difference in the number of species between areas upstream and downstream of a tributary is 1.84. The 95% confidence interval is -4.59 and 8.25. 

### Question B

> Test the hypothesis that the tributaries have no effect on the number of species of electric fish.

ANSWER

```{r}
t.test(formula = species ~ location, data = fish_long)
```

The p-value is more than 0.05, which means we retain the null hypothesis and reject the alternative. This means that the true difference in means is equal to 0 and that the tributaries have no effect on the number of species of electric fish. 

### Question C

> State the assumptions that you had to make to complete parts (A) and (B). Create a graph to assess whether one of those assumptions was met.

ANSWER

The assumptions that I made was using the p-value from the t-test and also making sure that there was a normal distribution for the graph. When plotting the histogram, I can tell that it is a normal distribution. 

```{r}
fish_long %>% 
  ggplot(aes(x = species)) +
  geom_histogram(
    aes(fill = location), 
    bins = 8, 
    alpha = 0.5, 
    position = "identity"
  ) +
  scale_fill_manual(values = c("darkorange","cyan4")) +
  theme_minimal()
```

## ANOVA

Fiddler crabs are so called because males have a greatly enlarged "major" claw, which is used to attract females and to defend a burrow.

Darnell and Munguia (2011) recently suggested that this appendage might also act as a heat sink, keeping males cooler while out of the burrow on hot days.

To test this, they placed four groups of crabs into separate plastic cups and supplied a source of radiant heat (60-watt light bulb) from above. The four groups were intact male crabs, male crabs with the major claw removed; male crabs with the other (minor) claw removed (control); and intact female fiddler crabs.

They measured the body temperature of crabs every 10 minutes for 1.5 hours. These measurements were used to calculate a rate of heat gain for every individual crab in degrees

### Question D

Graph the distribution of body temperatures for each crab type:

```{r}
crabs %>% 
  ggplot(aes(x = bodyTemperature)) +
  geom_histogram(
    aes(fill = crabType), 
    bins = 12, 
    alpha = 0.5, 
    position = "identity",
    na.rm = TRUE
  ) +
  theme_minimal()
```

### Question E

Does body temperature varies among crab types? State the null and alternative hypothesis, conduct and ANOVA, and interpret the results.

```{r}
aov_crab_temp <-
  aov(bodyTemperature ~ crabType, data = crabs)
aov_crab_temp
summary(aov_crab_temp)
```

The null hypothesis is the mean body temperatures for at least one crab type is not different from the means for the other crab types. The alternative hypothesis is the mean body temperatures for at least one crab type is different from the means for the other crab types. Since the p-value is way below 0.05 or below any other alpha we choose, that means that we will reject the null hypothesis and conclude that the mean body temperature for at least one crab type is different from the means for the other crab types. 